---
title: "Exploratory 1"
author: "Maisie Sun"
date: "2022-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(janitor)
library(ggridges)
library(patchwork)
library(readxl)
library("writexl")
library(rvest)
library(httr)
library(dplyr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Maternal Age
Get the data for age

```{r data_import_age}
url = "https://www.health.ny.gov/statistics/vital_statistics/2019/table21.htm"

induced_abortion = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

```

data cleaning for age

```{r data_cleaning_age}
clean_age = 
  induced_abortion %>%
  select(1:8) %>%
  purrr::set_names(c("borough", "total", "age_less_20", "age_20_24", "age_25_29", "age_30_34", "age_35_39", "age_plus_40")) %>%
  slice(6:11) %>%
  mutate(
    borough = str_replace(borough, "Kings", "Brooklyn"),
    borough = str_replace(borough, "New York City", "New_York_City"),
    borough = str_replace(borough, "New York", "Manhattan"),
    borough = str_replace(borough, "Richmond", "Staten_Island"),
    total = str_replace(total, ",", ""),
    age_less_20 = str_replace(age_less_20, ",", ""),
    age_20_24 = str_replace(age_20_24, ",", ""),
    age_25_29 = str_replace(age_25_29, ",", ""),
    age_30_34 = str_replace(age_30_34, ",", ""),
    age_35_39 = str_replace(age_35_39, ",", ""),
    age_plus_40 = str_replace(age_plus_40, ",", "")
    ) %>%
  mutate_at(c("total", "age_less_20", "age_20_24", "age_25_29", "age_30_34", "age_35_39", "age_plus_40"), as.numeric)
  

```

ggplot: induced abortion in boroughs

```{r total_abortion_plot}
total_abortion = 
  clean_age %>%
  slice(-1) %>%
  ggplot((aes(x = borough, y = total))) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  labs(
    title = "Number of Induced Abortions by Borough, 2019",
    x = "New York City Borough",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_total.pdf", total_abortion, width = 10, height = 5)

```

ggplot: induced abortion vs maternal age

```{r age_plot}
total_abortion_age = 
  clean_age %>%
  slice_head(n = 1) %>%
  select(-total) %>%
  pivot_longer(
    age_less_20:age_plus_40,
    names_to = "age", 
    values_to = "abortion"
  )

abortion_age_wtotal_plot = 
  clean_age %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    age_less_20:age_plus_40,
    names_to = "age", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = age, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_point(data = total_abortion_age) +
  labs(
    title = "Number of Induced Abortions by Woman's Age and Borough, 2019",
    x = "Woman's age categories",
    y = "Number of Induced Abortions"
  )

abortion_age_plot = 
  clean_age %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    age_less_20:age_plus_40,
    names_to = "age", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = age, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Number of Induced Abortions by Woman's Age and Borough, 2019",
    x = "Woman's age categories",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_age.pdf", abortion_age_plot, width = 10, height = 5)

ggsave("results/abortion_age_wtotal.pdf", abortion_age_wtotal_plot, width = 10, height = 5)

```

## Financial Plans

Get the data

```{r data_import_finance}
url = "https://www.health.ny.gov/statistics/vital_statistics/2019/table24.htm"

induced_abortion_finance = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

```

data cleaning

```{r data_cleaning_finance}
clean_finance = 
  induced_abortion_finance %>%
  select(1:8) %>%
  purrr::set_names(c("borough", "total", "Medicaid", "HMO", "Other_Insurance", "Self_Pay", "Multiple_Payor", "Not_Stated")) %>%
  slice(6:11) %>%
  mutate(
    borough = str_replace(borough, "Kings", "Brooklyn"),
    borough = str_replace(borough, "New York City", "New_York_City"),
    borough = str_replace(borough, "New York", "Manhattan"),
    borough = str_replace(borough, "Richmond", "Staten_Island"),
    total = str_replace(total, ",", ""),
    Medicaid = str_replace(Medicaid, ",", ""),
    HMO = str_replace(HMO, ",", ""),
    Other_Insurance = str_replace(Other_Insurance, ",", ""),
    Self_Pay = str_replace(Self_Pay, ",", ""),
    Multiple_Payor = str_replace(Multiple_Payor, ",", ""),
    Not_Stated = str_replace(Not_Stated, ",", "")
    ) %>%
  mutate_at(c("total", "Medicaid", "HMO", "Other_Insurance", "Self_Pay", "Multiple_Payor", "Not_Stated"), as.numeric)
  

```

ggplot: induced abortion vs financial plans

```{r viz_finance}
total_abortion_finance = 
  clean_finance %>%
  slice_head(n = 1) %>%
  select(-total) %>%
  pivot_longer(
    Medicaid:Not_Stated,
    names_to = "finance_plan", 
    values_to = "abortion"
  )

abortion_finance_plot = 
  clean_finance %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    Medicaid:Not_Stated,
    names_to = "finance_plan", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = finance_plan, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Number of Induced Abortions by Woman's Finance Plan and Borough, 2019",
    x = "Woman's finance plan",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_finance_plan.pdf", abortion_finance_plot, width = 10, height = 5)

abortion_finance_total = 
  clean_finance %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    Medicaid:Not_Stated,
    names_to = "finance_plan", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = finance_plan, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_point(data = total_abortion_finance) +
  labs(
    title = "Number of Induced Abortions by Woman's Finance Plan and Borough, 2019",
    x = "Woman's finance plan",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_finance_total.pdf", abortion_finance_total, width = 10, height = 5)
```

## Race

```{r data_import_race}
url = "https://www.health.ny.gov/statistics/vital_statistics/2019/table23.htm"

induced_abortion = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

```

data cleaning

```{r data_cleaning_race}
clean_nh = 
  induced_abortion %>%
  select(1,2,4,6,8,10) %>% 
  purrr::set_names(c("borough", "total", "Non_Hispanic_Total", "NH_White_Only_Number", "NH_Black_Only_Number", "Other_Number")) %>% 
    slice(6:11) %>%
  mutate(
    borough = str_replace(borough, "Kings", "Brooklyn"),
    borough = str_replace(borough, "New York City", "New_York_City"),
    borough = str_replace(borough, "New York", "Manhattan"),
    borough = str_replace(borough, "Richmond", "Staten_Island"),
    total = str_replace(total, ",", ""),
    Non_Hispanic_Total = str_replace(Non_Hispanic_Total, ",", ""),
    NH_White_Only_Number = str_replace(NH_White_Only_Number, ",", ""),
    NH_Black_Only_Number = str_replace(NH_Black_Only_Number, ",", ""),
    Other_Number = str_replace(Other_Number, ",", ""),
    ) %>%
  mutate_at(c("total", "Non_Hispanic_Total", "NH_White_Only_Number", "NH_Black_Only_Number", "Other_Number"), as.numeric)
```

ggplot: induced abortion vs race (Non-Hispanic)

```{r viz_race}
total_abortion_nhrace = 
  clean_nh %>%
  slice_head(n = 1) %>%
  pivot_longer(
    total:Other_Number,
    names_to = "race", 
    values_to = "abortion"
  )

abortion_race_nhplot = 
  clean_nh %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    Non_Hispanic_Total:Other_Number,
    names_to = "race", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = race, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Number of Induced Abortions by Race (Non-Hispanic) and Borough, 2019",
    x = "Woman's Race",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_nhrace.pdf", abortion_race_nhplot, width = 10, height = 5)

abortion_race_nhtotal = 
  clean_nh %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    Non_Hispanic_Total:Other_Number,
    names_to = "race", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = race, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_point(data = total_abortion_nhrace) +
  labs(
    title = "Number of Induced Abortions by Race (Non-Hispanic) and Borough, 2019",
    x = "Woman's Finance Plan",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_race_nhtotal.pdf", abortion_race_nhtotal, width = 10, height = 5)
```

data cleaning (Hispanic)

```{r data_cleaning_race}
clean_h = 
  induced_abortion %>%
  select(1,2,12,14,16) %>% 
  purrr::set_names(c("borough", "total", "Hispanic_Total", "H_White_Only_Number", "H_Black_Only_Number")) %>% 
    slice(6:11) %>%
  mutate(
    borough = str_replace(borough, "Kings", "Brooklyn"),
    borough = str_replace(borough, "New York City", "New_York_City"),
    borough = str_replace(borough, "New York", "Manhattan"),
    borough = str_replace(borough, "Richmond", "Staten_Island"),
    total = str_replace(total, ",", ""),
    Hispanic_Total = str_replace(Hispanic_Total, ",", ""),
    H_White_Only_Number = str_replace(H_White_Only_Number, ",", ""),
    H_Black_Only_Number = str_replace(H_Black_Only_Number, ",", "")
    ) %>%
  mutate_at(c("total", "Hispanic_Total", "H_White_Only_Number", "H_Black_Only_Number"), as.numeric)
```

ggplot: induced abortion vs race (Hispanic)

```{r viz_race}
total_abortion_hrace = 
  clean_h %>%
  slice_head(n = 1) %>%
  pivot_longer(
    Hispanic_Total:H_Black_Only_Number,
    names_to = "race", 
    values_to = "abortion"
  )

abortion_race_h = 
  clean_h %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    Hispanic_Total:H_Black_Only_Number,
    names_to = "race", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = race, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Number of Induced Abortions by Race (Hispanic) and Borough, 2019",
    x = "Woman's Race (Non-Hispanic)",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_hrace.pdf", abortion_race_h, width = 10, height = 5)

abortion_race_htotal = 
  clean_h %>%
  select(-total) %>%
  slice(-1) %>%
  pivot_longer(
    Hispanic_Total:H_Black_Only_Number,
    names_to = "race", 
    values_to = "abortion"
  ) %>%
  ggplot(aes(x = race, y = abortion, fill = borough)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_point(data = total_abortion_hrace) +
  labs(
    title = "Number of Induced Abortions by Race (Hispanic) and Borough, 2019",
    x = "Woman's Race (Hispanic)",
    y = "Number of Induced Abortions"
  )

ggsave("results/abortion_race_htotal.pdf", abortion_race_htotal, width = 10, height = 5)
```

## Combine tables and clean the table

```{r}
abortion_data = 
  full_join(clean_h, clean_nh)
abortion_data = 
  full_join(abortion_data, clean_finance)
abortion_data = 
  full_join(abortion_data, clean_age)

abortion_data = 
  abortion_data %>%
  t() %>%
  as.data.frame() %>%
  purrr::set_names(c("New_York_City", "Bronx", "Brooklyn", "Manhattan", "Queens", "Staten_Island")) %>%
  slice(-1) %>%
  tibble::rownames_to_column()

write.csv(abortion_data, "data/abortion_data.csv", row.names = FALSE)

```


