---
title: "Abortion by Year"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
library(janitor)
library(ggridges)
library(patchwork)
library(readxl)
library("writexl")
library(rvest)
library(httr)
library(dplyr)
library(plotly)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Induced Abortions in Years (2019)

```{r data_import_2019}
url = "https://www.health.ny.gov/statistics/vital_statistics/2019/table22.htm"
url1= "https://www.health.ny.gov/statistics/vital_statistics/2019/table13.htm"

induced_abortion_2019 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

live_birth_2019 = 
  read_html(url1) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()
```

data cleaning

```{r data_cleaning_2019, echo = FALSE}
clean_ia_2019 = 
  induced_abortion_2019 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2019") %>%
  mutate_at(c("total", "year"), as.numeric)

clean_lb_2019 = 
  live_birth_2019  %>%  
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2019") %>%
  mutate_at(c("total", "year"), as.numeric)
```

#### Renaming

created regions based on this website: https://statejobs.ny.gov/assets/help/regionMapText.cfm

renaming counties into regions.

```{r rename}
rename_ia_2019= clean_ia_2019 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))

rename_lb_2019= clean_lb_2019 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex|Hamilton", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))
```

## Induced Abortions in Years (2018)
```{r data_import_2018}
url = "https://www.health.ny.gov/statistics/vital_statistics/2018/table22.htm"
url1 = "https://www.health.ny.gov/statistics/vital_statistics/2018/table07.htm"

induced_abortion_2018 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

live_birth_2018 = 
  read_html(url1) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()
```

data cleaning

```{r data_cleaning_2018, echo = FALSE}
clean_ia_2018 = 
  induced_abortion_2018 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2018"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)

clean_lb_2018 = 
  live_birth_2018 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(
    borough = str_replace(borough, "Kings", "Brooklyn"),
    borough = ifelse(as.character(borough) == "New York", "Manhattan", as.character(borough)),
    borough = str_replace(borough, "Richmond", "Staten_Island"),
    total = str_replace(total, ",", ""),
    year="2018"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)
```

#### Renaming

renaming counties into regions.

```{r rename}
rename_lb_2018= clean_lb_2018 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex|Hamilton", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))

rename_ia_2018= clean_ia_2018 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))
```

## Induced Abortions in Years (2017)
```{r data_import_2017}
url = "https://www.health.ny.gov/statistics/vital_statistics/2017/table22.htm"
url1= "https://www.health.ny.gov/statistics/vital_statistics/2017/table13.htm"

induced_abortion_2017 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

live_birth_2017 = 
  read_html(url1) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()
```

data cleaning

```{r data_cleaning_2017, echo = FALSE}
clean_ia_2017 = 
  induced_abortion_2017 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2017"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)

clean_lb_2017 = 
  live_birth_2017%>%  
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2019"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)
```

renaming counties into regions.

```{r rename live births}
rename_lb_2017= clean_lb_2017 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex|Hamilton", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))

rename_ia_2017= clean_ia_2017 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))
```

## Induced Abortions in Years (2016)
```{r data_import_2016}
url = "https://www.health.ny.gov/statistics/vital_statistics/2016/table21.htm"
url1= "https://www.health.ny.gov/statistics/vital_statistics/2016/table13.htm"

induced_abortion_2016= 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

live_birth_2016= 
  read_html(url1) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

```

data cleaning

```{r data_cleaning_2016, echo = FALSE}
clean_ia_2016 = 
  induced_abortion_2016 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2016"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)

clean_lb_2016 = 
  live_birth_2016 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2016") %>%
  mutate_at(c("total", "year"), as.numeric)
```

renaming counties into regions.

```{r rename live births}
rename_lb_2016= clean_lb_2016 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex|Hamilton", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))

rename_ia_2016= clean_ia_2016 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))
```

## Induced Abortions in Years (2015)
```{r data_import_2015}
url = "https://www.health.ny.gov/statistics/vital_statistics/2015/table21.htm"
url1 = "https://www.health.ny.gov/statistics/vital_statistics/2015/table07.htm"

induced_abortion_2015 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

live_birth_2015 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()
```

data cleaning

```{r data_cleaning_2015, echo = FALSE}
clean_ia_2015 = 
  induced_abortion_2015 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2015") %>%
  mutate_at(c("total", "year"), as.numeric)

clean_lb_2015 = 
  live_birth_2015 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2015") %>%
  mutate_at(c("total", "year"), as.numeric)
```

renaming counties into regions.

```{r rename live births}
rename_lb_2015= clean_lb_2015 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))

rename_ia_2015= clean_ia_2015 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))
```

## Induced Abortions in Years (2014)
```{r data_import_2014}
url = "https://www.health.ny.gov/statistics/vital_statistics/2014/table21.htm"
url1 = "https://www.health.ny.gov/statistics/vital_statistics/2014/table07.htm"

induced_abortion_2014 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

live_birth_2014 = 
  read_html(url) %>%
  html_table(header = FALSE) %>%
  first() %>%
  janitor::clean_names()

```

data cleaning

```{r data_cleaning_2014, echo = FALSE}
clean_ia_2014 = 
  induced_abortion_2014 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2014"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)

clean_lb_2014 = 
  live_birth_2014 %>%
  select(1:2) %>%
  purrr::set_names(c("borough", "total")) %>%
  slice(14:70) %>% 
  mutate(total = str_replace(total, ",", ""),
    year="2014"
    ) %>%
  mutate_at(c("total", "year"), as.numeric)
```

#### Renaming

renaming counties into regions.

```{r rename}
rename_lb_2014= clean_lb_2014 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))

rename_ia_2014= clean_ia_2014 %>% 
  transform(borough = gsub(pattern = "Albany|Columbia|Fulton|Greene|Montgomery|Rensselaer|Saratoga|Schenectady|Schoharie|Warren|Washington", replacement = "Saratoga", borough)) %>%
  transform(borough = gsub(pattern = "Franklin|Clinton|Essex/Hamilton|Hamilton/Essex", replacement = "Eastern Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Herkimer|Jefferson|Lewis|Oneida|Oswego|St Lawrence", replacement = "Western Adirondacks", borough)) %>%
  transform(borough = gsub(pattern = "Broome|Cayuga|Chenango|Cortland|Madison|Onondaga|Otsego|Tioga|Tompkins", replacement = "Central New York", borough)) %>% 
  transform(borough=gsub(pattern="Chemung|Genesee|Livingston|Monroe|Ontario|Schuyler|Seneca|Steuben|Wayne|Yates",replacement="Finger Lakes", borough)) %>% 
  transform(borough=gsub(pattern="Allegany|Cattaraugus|Chautauqua|Erie|Niagara|Orleans|Wyoming",replacement="Western New York", borough)) %>% 
  transform(borough=gsub(pattern="Delaware|Dutchess|Orange|Putnam|Sullivan|Ulster",replacement="Hudson Valley", borough)) %>% 
  transform(borough=gsub(pattern="Rockland|Westchester", replacement="Westchester/Rockland", borough)) %>% 
  transform(borough=gsub(pattern="Suffolk|Nassau", replacement="Long Island", borough)) %>% 
  group_by(borough, year) %>% 
  summarize(total_rate= sum(total))
```

Merge all datasets

```{r join}
merged_data=
bind_rows(rename_lb_2019, rename_ia_2019, rename_lb_2018, rename_ia_2018, by=rename_lb_2017, rename_ia_2017, rename_lb_2016, rename_ia_2016, rename_lb_2015, rename_ia_2015, rename_lb_2014, rename_ia_2014) %>%
select(year, everything()) %>% 
filter(!year=="year") %>% 
mutate_at(c("total_rate", "year"), as.numeric) 
```

#### Line chart by Year and Borough

```{r}
plot_borough_year=merged_data %>% 
  mutate(borough = fct_reorder(borough, total_rate)) %>% 
plot_ly(y = ~total_rate, x=~year, color = ~borough, type = "scatter", mode="line", colors = "viridis") %>% 
   layout(title = 'Induced Abortions Year by Borough', yaxis = list(title = 'Number of Induced Abortions per 1,000 Live Births'))

plot_borough_year
```
